---
title: "sims"
output: html_document
---
# Setup
#### Functions and libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(admixturegraph)
library(neldermead)
library(doParallel)
library(parallel)
library(foreach)
library(pbapply)
library(LEA)
library(cluster)
library(viridis)
# library(gtools)
source("plotting_funcs.R")
# source("http://membres-timc.imag.fr/Olivier.Francois/Conversion.R")
# source("http://membres-timc.imag.fr/Olivier.Francois/POPSutilities.R")
```


#### Functions 
```{r}
gendata<-function(migration_command,n,ninds){
  hapdat<-matrix(unlist(lapply(strsplit(system(paste("mspms ", n*ninds, " 1 -t 450 -r 450 450 -I ", n, " ", paste(rep(ninds,n),collapse=" ") , " ", migration_command, " | tail -", n*ninds,sep=""),intern=T),''),as.numeric)),nrow=n*ninds,byrow=T) # run coalescent sim
  
  genodat<-hapdat[seq(1,n*ninds,2),]+hapdat[seq(2,n*ninds,2),] # make diploid 
  
  write.table(t(genodat),file="~/crap.geno", quote=F, row.names=F, col.names=F, sep="") #write genotype file
  return(genodat)
}

'%!in%' <- function(x,y)!('%in%'(x,y)) # "not a member of"" function

poly<-function(x){ if(x>0 && x<1){ return(TRUE)} else{return(FALSE)}} # find polymorphic loci

f4_calc<-function(perms,list_of_freqs,locus){
  freqs<-sapply(perms,function(z) list_of_freqs[[z]][locus])
  f4_freqs=freqs[perm]
  return((f4_freqs[1]-f4_freqs[2])*(f4_freqs[3]-f4_freqs[4]))
} #calculate f4 stats (not used)
```

## Simulation  
#### Coalescent Simulation
```{r}
npops=6
ninds=20
kpops=6

#model w/ admixture but no migration
migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -es 0.07 6 0.8 -ej 0.07 6 3 -ej 0.07 7 2 -ej 0.1 3 1 -ej 0.6 2 1")

#model with admixture and migration 
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -m 4 3 50.0 -m 3 4 50.0 -m 2 6 50.0 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -es 0.07 6 0.8 -ej 0.07 6 3 -ej 0.07 7 2 -ej 0.1 3 1 -ej 0.6 2 1")

#model migration with no pop split
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -m 4 3 1.0 -m 3 4 1 -m 2 6 .5 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -ej 0.07 6 3 -ej 0.1 3 1 -ej 0.6 2 1")

#nothing
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -ej 0.07 6 3 -ej 0.1 3 1 -ej 0.6 2 1")

mydat<-gendata(migration_command,npops,ninds)
```

#### Set up admix graph
Set Tree
```{r set tree, fig.height=3.5, fig.width=4, results="hide"}
leaves <- c("SH", "SL", "MH", "ML","PARV","MEX")
inner_nodes <- c("S", "M", "SM","PM","PX")
edges <- parent_edges(c(edge("SH", "S"),
                        edge("SL", "S"),
                        edge("ML", "M"),
                     #  admixture_edge("oMH", "MM","M", "α"),
                        edge("MH", "M"),                          
                        edge("M","SM"),
                    #   edge("MH","oMH"),
                        edge("S","SM"),
                        edge("PARV","PM"),
                        edge("SM","PM"),
                        edge("PM","PX"),
                        edge("MEX","PX")
                        ))
graph <- agraph(leaves, inner_nodes, edges)
make_an_outgroup(graph, "MEx")
plot(graph, show_inner_node_labels = TRUE, show_admixture_labels = TRUE)
```

Make list of trees
```{r}
graph_list <- add_an_admixture(graph, "MIX")
```

#### Silhouette scores

Make distance matrix
```{r}
nloci=dim(mydat)[2]
ibs_mat<-as.matrix(dist(mydat,method="manhattan",diag=TRUE,upper=TRUE)/nloci)
pops=c("PARV","MEX","ML","SL","SH","MH")
for_sil=data.frame(paste(1:(npops*ninds/2),"@",c(sapply(pops,function(j) rep(j,ninds/2))),sep=""),ibs_mat)
write.table(for_sil,file="sim_dist_matrix.txt",col.names=FALSE,row.names=FALSE,quote=FALSE,sep="\t")
```

Calculate silhoutte scores, make graphs
```{r}
system("perl silhouette.pl -i sim_dist_matrix.txt -o temp -g")
```

Get individuals cut due to silhouette scores
```{r,message=FALSE}
#read in silscores
tempcut<-data.frame(read.table("temp_ind.sil",header=F),expand.grid(1:(ninds/2),1:npops))
colnames(tempcut)=c("pop","silscore","samplenum","popnum")

#get list of inds with silscores<0
cuts=sapply(1:npops, function(p) filter(tempcut,silscore<0 & popnum==p)$samplenum)

#cut based on silhouette scores
keep<-sort(which(silhouette(kmeans(dist(mydat),npops)[[1]],dist(mydat))[,3]>0))

left_out<-which(!(1:(npops*ninds/2) %in% keep))
cut_out<-paste(left_out,tempcut$pop[left_out])

if(length(left_out)>0){ 
  print(paste("The following",length(left_out),"individuals were removed:"))
  print(cut_out)
} else {
  print("Nobody cut.")
}

#make new data from cut
cutdat<-mydat[keep,] 
write.table(t(cutdat),file="~/crapcut.geno",quote=F,row.names=F,col.names=F,sep="")
```

#### Allele counts
Allele counts without silhouette cuts
```{r}
allele_counts=list()
allele_freqs=list()
cumulative=0
rmpop=as.numeric()
for(i in 1:(npops)){
  sample=1:(ninds/2)
  if(length(sample)==0){  
      allele_counts[[i]]=paste(0,0,sep=",")
      allele_freqs[[i]]=paste(0,0,sep=",")
      rmpop=c(rmpop,i)
  } else if(length(sample)==1) {
    allele_counts[[i]]=paste(mydat[cumulative+sample,],2*length(sample)-mydat[cumulative+sample,],sep=",")
    allele_freqs[[i]]=mydat[cumulative+sample,]/(2*length(sample))
  }
  else{
    allele_counts[[i]]=paste(colSums(mydat[cumulative+sample,]),2*length(sample)-colSums(mydat[cumulative+sample,]),sep=",")
    allele_freqs[[i]]=colSums(mydat[cumulative+sample,])/(2*length(sample))
  }
  cumulative=cumulative+max(sample)
}
```

Allele counts with silhouette cuts
```{r}
cut_allele_counts=list()
cut_allele_freqs=list()
cum=0
rmpop=as.numeric()
if(length(left_out)==0){ 
  cut_allele_counts=allele_counts
  cut_allele_freqs=allele_freqs
} else{
  for(i in 1:(npops)){
    sample=which(1:(ninds/2) %!in% cuts[[i]])
    if(length(sample)==0){  
      cut_allele_counts[[i]]=paste(0,0,sep=",")
      cut_allele_freqs[[i]]=paste(0,0,sep=",")
      rmpop=c(rmpop,i)
    } else if(length(sample)==1) {
      cut_allele_counts[[i]]=paste(mydat[cum+sample,],2*length(sample)-mydat[cum+sample,],sep=",")
      cut_allele_freqs[[i]]=mydat[cum+sample,]/(2*length(sample))
    }
    else{
      cut_allele_counts[[i]]=paste(colSums(mydat[cum+sample,]),2*length(sample)-colSums(mydat[cum+sample,]),sep=",")
      cut_allele_freqs[[i]]=colSums(mydat[cum+sample,])/(2*length(sample))
    }
    cum=cum+max(sample)
  }
}

#test: 
#sapply(1:6, function(p) sum(allele_counts[[p]]!=cut_allele_counts[[p]]))
```


## Structure Plots

#### Uncut
```{r}
#structure on original
obj.snmf<-snmf("~/crap.geno",K = kpops, alpha = 10, project = "new") 
qmatrix = Q(obj.snmf, K = kpops)
bobuncut=gather(data.frame(t(qmatrix)),dude,qval,1:(nrow(qmatrix))) %>%
mutate(pop=factor(rep(1:kpops,nrow(qmatrix))),dude=factor(as.numeric(gsub("X","",dude))))
```

#### Cut
```{r, include=FALSE}
#structure on cut data
obj.snmf<-snmf("~/crapcut.geno",K = kpops, alpha = 10, project = "new") 
qmatrix = Q(obj.snmf, K = kpops)

#make new matrix with 0s for left out individuals
if(length(left_out)>0){
  m <- matrix(0, ncol=kpops,nrow=(npops*ninds/2)) 
  m[!(1:nrow(m) %in% left_out),] <- qmatrix
} else{ m=qmatrix} 

bobcut=gather(data.frame(t(m)),dude,qval,1:(nrow(m))) %>%
  mutate(pop=factor(rep(1:kpops,nrow(m))),dude=factor(as.numeric(gsub("X","",dude))))
```

#### Plots
```{r,fig.height=3, fig.width=12, results="hide"}
uncut<-ggplot(bobuncut,aes(x=dude,y=qval,fill=pop))+
    geom_bar(stat="identity")+
    theme(
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank()
    )+
    ggtitle("OG data")+
    scale_fill_manual(guide=FALSE,values=viridis(npops))+
    scale_x_discrete(labels=1:(npops*ninds/2))

cut<-ggplot(bobcut,aes(x=dude,y=qval,fill=pop))+
  geom_bar(stat="identity")+
  ggtitle("silhouette cuts")+
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(),
        axis.title.x=element_blank(), axis.title.y=element_blank())+
  scale_fill_manual(guide=FALSE,values=viridis(npops)) 


bob=list(cut,uncut)
graphlist <- replicate(10, qplot(1,1), simplify = FALSE)

#title <- ggdraw() + draw_label(paste(migration_command), fontface='bold')

plots<-plot_grid(uncut,cut)
print( plot_grid(plots,ncol=1, rel_heights=c(0.3, 1)))
```


## Treemix
#### Uncut
```{r}
treemix<-data.frame(matrix(unlist(allele_counts),ncol=npops,byrow=F))
colnames(treemix)=c("PARV","MEX","ML","SL","SH","MH")
write.table(treemix,"treemix_input.txt",quote=F,row.name=F)
system("gzip -f treemix_input.txt")
```

Make tree
```{r}
system("~/src/treemix/src/treemix -i ~/Desktop/sandbox/silhouette/treemix_input.txt.gz -o treemix_out -m 1 -root MEX")
plot_tree("~/Desktop/sandbox/silhouette/treemix_out")
```

#### Cut
```{r}
treemix<-data.frame(matrix(unlist(cut_allele_counts),ncol=npops,byrow=F))
colnames(treemix)=c("PARV","MEX","ML","SL","SH","MH")
write.table(treemix,"treemix_cut_input.txt",quote=F,row.name=F)
system("gzip -f treemix_cut_input.txt")
```

Make tree
```{r}
system("~/src/treemix/src/treemix -i ~/Desktop/sandbox/silhouette/treemix_cut_input.txt.gz -o treemix_cut_out -m 1 -root MEX")
plot_tree("~/Desktop/sandbox/silhouette/treemix_cut_out")
```

#### Admixture Graphs
Get Data
```{r OG_Data}
system("~/src/treemix/src/fourpop -i ~/Desktop/sandbox/silhouette/treemix_input.txt.gz -k 100 | grep -Ev 'Estimating|total|npop' | sed -e 's/,/ /g;s/;/ /' > ./for_admix.txt")
OG<-read.table("./for_admix.txt",header=F) 
colnames(OG)=c("W","X","Y","Z","D","seD","Z.value")
OG <- OG[,c(1:5,7)]
plot(f4stats(OG))
```

Silhouette cuts
```{r}
system("~/src/treemix/src/fourpop -i ~/Desktop/sandbox/silhouette/treemix_cut_input.txt.gz -k 100 | grep -Ev 'Estimating|total|npop' | sed -e 's/,/ /g;s/;/ /' > ./for_admix_cut.txt")
cutdat<-read.table("./for_admix_cut.txt",header=F) 
colnames(cutdat)=c("W","X","Y","Z","D","seD","Z.value")
cutdat <- cutdat[,c(1:5,7)]
plot(f4stats(cutdat))
```


Fit graphs, plot graph with lowest error. First on OG data:
```{r}
ncores=3 #cpu to use
fits=fit_graph_list(OG,graph_list,ncores)
fits_errors=pbsapply(1:length(fits), function(i) (fits[[i]]$best_error))
plot(fits[[which(fits_errors==min(fits_errors))]]$graph)
plot(fit_graph(OG,fits[[which(fits_errors==min(fits_errors))]]$graph))
plot(fit_graph(OG,graph))
```

Then on silhouette scores:
```{r}
ncores=3 #cpu to use
fits=fit_graph_list(OG,graph_list,ncores)
fits_errors=pbsapply(1:length(fits), function(i) (fits[[i]]$best_error))
plot(fits[[which(fits_errors==min(fits_errors))]]$graph)
plot(fit_graph(OG,fits[[which(fits_errors==min(fits_errors))]]$graph))
plot(fit_graph(OG,graph))
```


