---
title: "sims"
output: html_document
---

## Functions and libraries
Libraries
```{r}
library(tidyverse)
library(admixturegraph)
source("plotting_funcs.R")
library(admixturegraph)
library(neldermead)
library(doParallel)
library(parallel)
library(foreach)
```


Functions to generate diploid data, remove monomorphic loci
```{r}
gendata<-function(migration_command,n,ninds){
  hapdat<-matrix(unlist(lapply(strsplit(system(paste("mspms ", n*ninds, " 1 -t 450 -r 450 450 -I ", n, " ", paste(rep(ninds,n),collapse=" ") , " ", migration_command, " | tail -", n*ninds,sep=""),intern=T),''),as.numeric)),nrow=n*ninds,byrow=T)
  
  genodat<-hapdat[seq(1,n*ninds,2),]+hapdat[seq(2,n*ninds,2),] # make diploid 
  
  write.table(t(genodat),file="~/crap.geno", quote=F, row.names=F, col.names=F, sep="") #write genotype file
  return(genodat)
}

'%!in%' <- function(x,y)!('%in%'(x,y)) # "not a member of"" function

poly<-function(x){ if(x>0 && x<1){ return(TRUE)} else{return(FALSE)}} # find polymorphic loci

f4_calc<-function(perms,list_of_freqs,locus){
  freqs<-sapply(perms,function(z) list_of_freqs[[z]][locus])
  f4_freqs=freqs[perm]
  return((f4_freqs[1]-f4_freqs[2])*(f4_freqs[3]-f4_freqs[4]))
}
```

## Simulation
#### Coalescent Simulation
```{r, message=FALSE, warning=FALSE, results='hide'}
npops=6
ninds=20
kpops=6

#model w/ admixture but no migration
migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -es 0.07 6 0.8 -ej 0.07 6 3 -ej 0.07 7 2 -ej 0.1 3 1 -ej 0.6 2 1")

#model with admixutre and migration 
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -m 4 3 50.0 -m 3 4 50.0 -m 2 6 50.0 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -es 0.07 6 0.8 -ej 0.07 6 3 -ej 0.07 7 2 -ej 0.1 3 1 -ej 0.6 2 1")

#model migration with no pop split
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -m 4 3 50.0 -m 3 4 50.0 -m 2 6 100.0 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -ej 0.07 6 3 -ej 0.1 3 1 -ej 0.6 2 1")

#nothing
#migration_command=paste("-n 4 0.8 -n 5 0.1 -n 6 0.8 -g 3 5.0 -ej 0.04 5 4 -ej 0.06 4 3 -ej 0.07 6 3 -ej 0.1 3 1 -ej 0.6 2 1")

mydat<-gendata(migration_command,npops,ninds)
```

#### Set up admix graph
##### Set Tree
```{r}
leaves <- c("SH", "SL", "MH", "ML","PARV","MEX")
inner_nodes <- c("S", "M", "SM","PM","PX","MM")
edges <- parent_edges(c(edge("SH", "S"),
                        edge("SL", "S"),
                        edge("ML", "M"),
                     #  admixture_edge("oMH", "MM","M", "α"),
                        edge("MH", "M"),                                    
                        edge("M","SM"),
                    #   edge("MH","oMH"),
                        edge("S","SM"),
                        edge("PARV","PM"),
                        edge("SM","PM"),
                        edge("PM","PX"),
                        edge("MM","PX"),
                       edge("MEX","MM")
                        ))
graph <- agraph(leaves, inner_nodes, edges)
plot(graph, show_inner_node_labels = TRUE, show_admixture_labels = TRUE)
```

##### Make list of trees
```{r}
graph_list <- add_an_admixture2(graph, "MIX")
```

## OG Data

#### Allele counts
Allele counts without silhouette cuts
```{r}
allele_counts=list()
allele_freqs=list()
cumulative=0
rmpop=as.numeric()
for(i in 1:(npops)){
  sample=1:(ninds/2)
  if(length(sample)==0){  
      allele_counts[[i]]=paste(0,0,sep=",")
      allele_freqs[[i]]=paste(0,0,sep=",")
      rmpop=c(rmpop,i)
  } else if(length(sample)==1) {
    allele_counts[[i]]=paste(mydat[cumulative+sample,],2*length(sample)-mydat[cumulative+sample,],sep=",")
    allele_freqs[[i]]=mydat[cumulative+sample,]/(2*length(sample))
  }
  else{
    allele_counts[[i]]=paste(colSums(mydat[cumulative+sample,]),2*length(sample)-colSums(mydat[cumulative+sample,]),sep=",")
    allele_freqs[[i]]=colSums(mydat[cumulative+sample,])/(2*length(sample))
  }
  cumulative=cumulative+max(sample)
}
```

#### Treemix
```{r}
treemix<-data.frame(matrix(unlist(allele_counts),ncol=npops,byrow=F))
#colnames(treemix)=paste("pop",rep(1:(npops)),sep="")
colnames(treemix)=c("PARV","MEX","ML","SL","SH","MH")
write.table(treemix,"treemix_input.txt",quote=F,row.name=F)
system("gzip -f treemix_input.txt")
```

Make tree
```{r}
system("~/src/treemix/src/treemix -i ~/Desktop/sandbox/silhouette/treemix_input.txt.gz -o treemix_out -m 1")
plot_tree("~/Desktop/sandbox/silhouette/treemix_out")
```

#### Structure Plots
```{r}
#TODO
```

#### Admixture Graphs
Get Data
```{r OG_Data}
system("~/src/treemix/src/fourpop -i ~/Desktop/sandbox/silhouette/treemix_input.txt.gz -k 100 | grep -Ev 'Estimating|total|npop' | sed -e 's/,/ /g;s/;/ /' > ./for_admix.txt")
OG<-read.table("./for_admix.txt",header=F) 
colnames(OG)=c("W","X","Y","Z","D","seD","Z.value")
OG <- OG[,c(1:5,7)]
plot(f4stats(OG))
```

Fit graphs
```{r}
bob=fit_graph_list(bears,example_list_2,2)
ted=sapply(1:length(bob), function(i) (bob[[i]]$best_error))
plot(bob[[which(ted==min(ted))]]$graph)
```


## Silhouette

#### Calculate scores
Make distance matrix
```{r}
nloci=dim(mydat)[2]
ibs_mat<-as.matrix(dist(mydat,method="manhattan",diag=TRUE,upper=TRUE)/nloci)
pops=c("PARV","MEX","ML","SL","SH","MH")
for_sil=data.frame(paste(1:60,"@",c(sapply(pops,function(j) rep(j,10))),sep=""),ibs_mat)
write.table(for_sil,file="sim_dist_matrix.txt",col.names=FALSE,row.names=FALSE,quote=FALSE,sep="\t")
```

Calculate silhoutte scores, make graphs
```{r, message=FALSE, warning=TRUE, include=FALSE}
system("perl silhouette.pl -i sim_dist_matrix.txt -o temp -g")
```

Get individuals cut due to silhouettde scores
```{r}
#read in silscores
tempcut<-data.frame(read.table("temp_ind.sil",header=F),expand.grid(1:(ninds/2),1:npops))
colnames(tempcut)=c("pop","silscore","samplenum","popnum")

#get list of inds with silscores<0
cuts=sapply(1:npops, function(p) filter(tempcut,silscore<0 & popnum==p)$samplenum)
```

#### Allele Counts
Allele counts with silhouette cuts
```{r}
allele_counts=list()
allele_freqs=list()
cum=0
rmpop=as.numeric()
for(i in 1:(npops)){
  sample=which(1:(ninds/2) %!in% cuts[[i]])
  if(length(sample)==0){  
      allele_counts[[i]]=paste(0,0,sep=",")
      allele_freqs[[i]]=paste(0,0,sep=",")
      rmpop=c(rmpop,i)
  } else if(length(sample)==1) {
    allele_counts[[i]]=paste(mydat[cum+sample,],2*length(sample)-mydat[cum+sample,],sep=",")
    allele_freqs[[i]]=mydat[cum+sample,]/(2*length(sample))
  }
  else{
    allele_counts[[i]]=paste(colSums(mydat[cum+sample,]),2*length(sample)-colSums(mydat[cum+sample,]),sep=",")
    allele_freqs[[i]]=colSums(mydat[cum+sample,])/(2*length(sample))
  }
  cum=cum+max(sample)
}
```



